Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-06-17 paper.qmd
Added high-d projection

error_plot_two_curvy <- error_df_two_curvy_abs |>
  ggplot(aes(x = emb1,
             y = emb2,
             colour = sqrt_row_wise_total_error)) +
  geom_point(alpha=0.5) +
  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
  xlab("tSNE1") +
  ylab("tSNE2") +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )

error_plot_two_curvy_quasi <- ggplot(
  error_df_two_curvy_abs, 
  aes(x = sqrt_row_wise_total_error, 
      y = 0, 
      colour = sqrt_row_wise_total_error)) +
  geom_quasirandom(alpha = 0.5) +
  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
  xlab(expression(e[hj])) +
  ylab("") +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none")
```

@@

@@

<!-- Sparse end with dense corner to me is very hard to understand - sparse and dense are pretty much antonyms -->
```{r}
#| label: fig-p-d-error-in-2d-two-curvy
#| fig-cap: "The \\fD{} model error in \\gD{} layout. Color indicates error ($e_{hj}$), dark color indicating high error and light indicates low error. Most large errors are distributed near the sparse end of the nonlinear cluster and most small errors are distributed near the dense corners."
#| out-width: 80%
#| fig-pos: H

error_plot_two_curvy + error_plot_two_curvy_quasi +
  plot_layout(ncol=2)
```

>>>>

error_plot_two_curvy <- error_df_two_curvy_abs |>
  ggplot(aes(x = emb1,
             y = emb2,
             colour = error_cat)) +
  geom_point(alpha=0.5) +
  scale_colour_manual(values = c('#ffffcc','#ffeda0','#fed976',
                                 '#feb24c','#fd8d3c','#fc4e2a',
                                 '#e31a1c','#bd0026','#800026')) +
  xlab("tSNE1") +
  ylab("tSNE2") +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    legend.position = "none"
  )

error_plot_two_curvy_quasi <- ggplot(
  error_df_two_curvy_abs, 
  aes(x = sqrt_row_wise_total_error, 
      y = 0, 
      colour = error_cat)) +
  geom_quasirandom(alpha = 0.5) +
  scale_colour_manual(values = c('#ffffcc','#ffeda0','#fed976',
                                 '#feb24c','#fd8d3c','#fc4e2a',
                                 '#e31a1c','#bd0026','#800026')) +
  xlab(expression(e[hj])) +
  ylab("") +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none",
        plot.margin = margin(0, 0, 0, 0))
```

@@

projected_df <- proj_obj1$projected_df
axes <- proj_obj1$axes
circle <- proj_obj1$circle
cluster <- proj_obj1$cluster

projected_df <- projected_df |>
      dplyr::mutate(cluster = proj_obj1$cluster)

two_curvy_proj_tsne_first_model1_coloured <- projected_df |>
  ggplot(
    aes(
      x = proj1,
      y = proj2,
      colour = cluster)) +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2),
    label=rownames(axes),
    colour="grey50",
    size = 2) +
  geom_path(
    data=circle,
    aes(x=c1, y=c2), colour="grey70") +
  scale_colour_manual(values = c('#ffffcc','#ffeda0','#fed976',
                                 '#feb24c','#fd8d3c','#fc4e2a',
                                 '#e31a1c','#bd0026','#800026')) +
  xlim(c(-0.6, 0.6)) +
  ylim(c(-0.6, 0.6)) +
  theme(aspect.ratio = 1,
        legend.position = "none")

@@

```{r}
#| label: fig-p-d-error-in-2d-two-curvy
#| fig-cap: "The \\fD{} model error in \\gD{} layout. Color indicates error ($e_{hj}$), dark color indicating high error and light indicates low error. Most large errors are distributed near the sparse end of the nonlinear cluster and most small errors are distributed near the dense corners."
#| out-width: 80%
#| fig-pos: H

 error_plot_two_curvy_quasi + error_plot_two_curvy + two_curvy_proj_tsne_first_model1_coloured +
  plot_layout(ncol=3)
```
